---
title: "CRIM 250 Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Assignment 1

### Problem 1 
Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.
```{r}
#Install the dataset package  
#install.packages("datasets")

#load in the library 
library(datasets)
```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?
```{r}
#Load in the dataset
USArrests

#rename the USArrests dataset
dat <- USArrests
```
__It is useful to rename the dataset because it is easier to work with. If the data set had a longer or more complicated name, it would be difficult and time consuming to type out the dataset name every time when you want to perform a function on the data. It also lets you keep track of your work if you make different versions of it so that it is not contaminated by changes that were meant to be on one version and not the other. This will let you replicate the work if changes were made to different versions.__

### Problem 2
Use this command to make the state names into a new variable called State. 
```{r, eval=TRUE}
dat$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.

List the variables contained in the dataset `USArrests`.

```{r, eval=TRUE}
#find the variables in the dataset
names(dat)
```
 
__The variables in the USArrests dataset are Murder, Assault, UrbanPop, Rape, and State.__ 
 
### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

__Murder is a quantitative variable from the DVB chapter since it is a count of numbers.__

What R Type of variable is it?

__Murder is a numeric variable from the R type since it is also a count of numbers and functions such as mean or median can be applied.__
        
### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

__The dataset contains arrest numbers for 4 types of crimes within all 50 states. Each state has a corresponding arrest rate for murder, assault, and rape, as well as the percent of urban population within the state. The numbers mean the number of arrests per 100,000 people in that state.__

### Problem 5

Draw a histogram of `Murder` with proper labels and title.
```{r}
hist(dat$Murder, main = "Histogram of Murder Arrests", xlab = "Number of Murder Arrests", ylab = "Frequency")
```
 
### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
#finding the summary statistics of the murder variable
summary(dat$Murder)
```
__The mean of the murder variable is 7.788 murders, while the median is 7.250 murders. The mean is the sum of all data values divided by the number of values, or the average of the values within the dataset. The median, however, is the middle value after all of the values are put in numerical order. When distributions are skewed or there are outliers in the data, the median is better to use than the mean because the mean changes with skew or outliers and the median is more robust. A quartile is each of three values in which the data can be distributed into even fourths. I think that R gives us the 1st quartile and 3rd quartile because they can be used to calculate the interquartile range, which tells us the interval where half of the values within the data set lie. The median is also the 2nd quartile, so the three groups are given that split the data in even fourths.__
 
### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
#make a histogram of assault
hist(dat$Assault, main = "Histogram of Assault Arrests", xlab = "Number of Assault Arrests", ylab = "Frequency")

#make a histogram of rape
hist(dat$Rape, main = "Histogram of Rape Arrests", xlab = "Number of Rape Arrests", ylab = "Frequency")

#plot all three histograms together
par(mfrow=c(3,1))
hist(dat$Murder, main = "Histogram of Murder Arrests", xlab = "Number of Murder Arrests", ylab = "Frequency")
hist(dat$Assault, main = "Histogram of Assault Arrests", xlab = "Number of Assault Arrests", ylab = "Frequency")
hist(dat$Rape, main = "Histogram of Rape Arrests", xlab = "Number of Rape Arrests", ylab = "Frequency")
```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

__Par can be used to set the parameters on a graph, making it easier to combine multiple graphs into one.__

What can you learn from plotting the histograms together?

__By plotting these histograms together, you can compare the arrest rates for each crime. Murders happen the least frequently and assaults happen the most frequently. You can also see that the histogram for murder is unimodal and skewed, the histogram for assaults is bimodal, and the histogram for rapes is also unimodal and skewed.__ 

### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

```{r, fig.width = 7.5, fig.height = 4}
#install and load in the maps and ggplot2 packages
library('maps') 
library('ggplot2') 
```

Run this code:
```{r, fig.width = 7.5, fig.height = 4}
ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```
What does this code do? Explain what each line is doing.

__The first line above the map code loads the library for both packages. The first line of the map code is creating a plot of murder. The ggplot function is telling R to create a plot, the "dat" is telling R which dataset to use in the plot, and the "aes(map_id=state, fill=Murder))" is telling R to plot each state and fill it with the corresponding murder arrest rate. The geom_map function creates shapes for a reference map, and the map=map_data("state") function is telling R to map the data that corresponds with the state variable that we created earlier. The expand_limits function makes sure that the entire map fits within the plot of the graph and the (x=map_data("state")$long, y=map_data("state")$lat) tells R which data to base the plot limits off of, which in this case is the state data.__


# Assignment 2

### Problem 1
Load in the data.
```{r}
dat <- read.csv(file = "Assignment2data copy.csv")
```

What are the dimensions of the dataset? 
```{r}
dim(dat)
```
__There are 171 columns and 7 rows (or 171 respondents and 7 questions)__

### Problem 2
Describe the variables in the dataset.
```{r}
names(dat)
```
__There are 7 variables in the dataset:__
__1. "mjage" (How old were you the first time you used marijuana or hashish?)__
__2. "cigage" (How old were you when you first started smoking cigarettes everyday?)__
__3. "iralcage" (How old were you when you first tried alcohol?)__
__4. "age2" (Recoded final edited age (since respondents had multiple chances to change their age throughout the survey))__
__5. "sexatract" (Sexual attraction)__
__6. "speakengl" (How well do you speak English)__
__7. "irsex" (Imputation revised gender)__

__Mjage, cigage, and iralcage are numeric variables (or quantitative) because the respondent gave an exact number for their answers for each variable and functions like mean can be applied to find the average age the respondents tried marijuana, cigarettes, and alcohol. Age2, sexatract, speakengl, and irsex are categorical variables because the respondents' answers were split into categories when coded. Because respondents had the opportunity to change their age throughout the interview, the age variable was calculated from the raw birth date and the final edited interview date, the age entered in the questionnaire roster (if it exists), and the pre-interview screener age because interviewees had the opportunity to change their age throughout the interview.__

What is this dataset about? Who collected the data, what kind of sample is it, and what was the purpose of generating the data?
__The dataset is a small sample from the National Survey of Drug Use and Health, which was conducted by the Center for Behavioral Health Statistics and Quality (CBHSQ, formerly the Office of Applied Studies) within the Substance Abuse and Mental Health Services Administration (SAMHSA) and is conducted by RTI International, Research Triangle Park, North Carolina. The survey was conducted through a computer assisted administration, and was changed from a strictly national design to a state-based sampling plan in 1999. The primary purpose of generating the data was to measure the prevalence and correlation of substance use and mental health issues in the United States, according to the NSDUH 2019 Codebook.__

### Problem 3: Age and gender
What is the age distribution of the sample like? Make sure you read the codebook to know what the variable values mean.
```{r}
hist(dat$age2, main = "Histogram of Recoded Final Edited Age Categories", xlab = "Recoded Final Edited Age Categories")
```
__The age variable is skewed to the left, with the majority of the values in the 14-16 bin (which means that the majority of the people in the dataset are 30 to 64 years old, according to the codebook). There are fewer people who were very young when filling out the survey, with only a few in the 4-6 and 6-8 bins (meaning that the respondents in these bins were 15-17 and 18-19, respectively).__

Do you think this age distribution representative of the US population? Why or why  not?
```{r}
min(dat$age2) #the youngest respondent was 15 (category 4)
max(dat$age2) #the oldest respondent was 65 years old or older (category 17)
```
__Yes, I think that this age distribution is representative of the US population. The survey was conducted in order to evaluate drug use, and I would think that people would not try drugs before turning 15 (the youngest respondent) and very few people would continue using drugs when they are 65 years old or older (the oldest respondent). Also, according to the codebook, the participants were randomly selected to complete the survey, which fulfills one requirement of a representative sample. It makes sense that the number of respondents increased when the respondents were 19 to 23 (8 to 12 categories) as those are the ages when most people would be exposed to marijuana, drugs, and alcohol for the first time, and that the number of respondents increased again when the respondents were 24 to 64 (categories 12 to 16), as those are the ages that people would continue to use marijuana, drugs, and alcohol as adults (especially after potentially being exposed in college). I also think that it makes sense that the majority of respondents were 34 to 64 years old, because older generations are more likely to have smoked cigarettes everyday before perception and knowledge changed.__

Is the sample balanced in terms of gender? If not, are there more females or males?
```{r}
table(dat$irsex, dat$age2)
```
__The sample is not balanced in terms of gender. There were 91 males included in the dataset and 80 females included.__

Use this code to draw a stacked bar plot to view the relationship between sex and age. What can you conclude from this plot?
```{r}
tab.agesex <- table(dat$irsex, dat$age2)
barplot(tab.agesex,
        main = "Stacked barchart",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.agesex),
        beside = FALSE) # Stacked bars (default)
```
__From this plot, I can conclude that more males answered the survey than females, as most age categories have a higher frequency of males than females (especially in categories 6 and 7). I can also conclude than more older males answered the survey than older females. As indicated in both the plot and the table of sex and age categories, the highest five categories (13, 14, 15, 16, and 17) had more male responses than females (1 more, 4 more, 4 more, 4 more, and 5 more, respectively).__

### Problem 4: Substance use
For which of the three substances included in the dataset (marijuana, alcohol, and cigarettes) do individuals tend to use the substance earlier?
```{r}
boxplot(dat$mjage, dat$cigage, dat$iralcage, main = "Substance Usage and Age",
        ylab="Age", xlab="Types of Substances", 
        names=c("Marijuana or Hashish","Cigarettes","Alcohol"))
```
__Based on the boxplot of substance usage and age, it looks like individuals tend to use alcohol earlier than marijuana / hashish or earlier than they started smoking cigarettes every day.__

```{r}
min(dat$mjage) 
min(dat$cigage) 
min(dat$iralcage) 
```
__Looking at the minimum values for each type of substance, the minimum recorded age for when an individual first tried marijuana / hashish was 7, the minimum recorded age for when an individual first started smoking cigarettes everyday was 10, and the minimum recorded age for when an individual first tried alcohol was 5. This also confirms that like individuals tend to use alcohol earlier than marijuana / hashish or earlier than they started smoking cigarettes every day.__

### Problem 5: Sexual attraction
What does the distribution of sexual attraction look like? Is this what you expected?
```{r}
dat.sexat <- subset(dat$sexatract, subset = dat$sexatract<7)

hist(dat.sexat, main = "Distribution of Sexual Attraction", xlab = "Sexual Attraction")
```
__The distribution of sexual attraction is skewed to the right, with the majority of respondents answering that they are only attracted to the opposite sex (category one). The next greatest frequency of response is that the respondent is mostly attracted to the opposite sex (category 2), and then followed by the respondent is equally attracted to males and females (category 3). This is exactly what I expected because I would think that an overwhelming majority of people would be straight as that is most common in the United States.__

What is the distribution of sexual attraction by gender?
```{r}
table(dat$irsex, dat$sexatract)

library(dplyr) 

dat$sexatract <- dat$sexatract %>% na_if(., "85")
dat$sexatract <- dat$sexatract %>% na_if(., "94")
dat$sexatract <- dat$sexatract %>% na_if(., "97")
dat$sexatract <- dat$sexatract %>% na_if(., "98")
dat$sexatract <- dat$sexatract %>% na_if(., "99")

dat.five <- data.frame(dat$sexatract, dat$irsex)

barplot <- barplot(table(dat.five$dat.irsex, dat.five$dat.sexatract),
        main = "Sexual Attraction and Gender",
        xlab = "Sexaul Attraction Category",
        ylab = "Count",
        border = "black",
        col = c("hotpink1", "blue"),
        ylim = c(0,171),
        )
barplot
```
__Looking at the stacked bar plot of sexual attraction and gender, it confirms what was seen in the table that more males than females responded that they are only attracted to the opposite sex (category one). It also confirms that more females than males answered that they are mostly attracted to the opposite sex (category 2) and are equally attracted to males and females (category 3).__

### Problem 6: English speaking
What does the distribution of English speaking look like in the sample? 
```{r}
hist(dat$speakengl, main = "Histogram of English Speakers", xlab = "How Well You Speak English (1 - lowest, 4 - highest)")
```
__The distribution of English speaking looks extremely skewed to the right, with the majority of respondents answering that they speak English very well (category one).__

```{r}
table(dat$speakengl)
```
__Looking at the table of respondents who answered the “Speak English” question,161 people answered that the speak English very well, eight answered that they speak English well, and only two answered that they speak English not well. This confirms the skew seen in the histogram.__

Is this what you might expect for a random sample of the US population?
__This is what I would expect a random sample of the US population to look like, as the majority of people in the US speak English and speak it very well. Even if there were people included in the random sample from places that do not speak as much English, like Texas or Miami, for example, the large amount of people who do speak English well would account for what is seen in the histogram.__

Are there more English speaker females or males?
```{r}
table(dat$irsex, dat$speakengl)

barplot(table(dat$irsex, dat$speakengl),
        main="English Speaking Level and Gender",
        xlab="English Speaking Level Category",
        ylab="Count",
        col=c("hotpink1", "mediumpurple1"),
        ylim = c(0,200)
)
```
__Looking at the stacked bar plot of English speaking level and gender, it looks like the number of males and females that answered that they spoke English very well was pretty close, will a few more males (exactly 11, according to the table). The plot also confirms that more men responded that they spoke English well than women, and that no men responded that they spoke English not well.__

# Exam 1

## Instructions

a. Create a folder in your computer (a good place would be under Crim 250, Exams). 

b. Download the dataset from the Canvas website (fatal-police-shootings-data.csv) onto that folder, and save your Exam 1.Rmd file in the same folder.

c. Download the README.md file. This is the codebook. 

d. Load the data into an R data frame.
```{r}
dat <- read.csv(file = "data/fatal-police-shootings-datacopy.csv")
```


### Problem 1 (10 points)

a. Describe the dataset. This is the source: https://github.com/washingtonpost/data-police-shootings . Write two sentences (max.) about this.

__The dataset is made up of every fatal police shooting in the US since January 1, 2015, as collected by the Washington Post. The dataset includes 6594 people and 17 variables, which contain important information about each shooting, such as the name of the person shot, their mental health status, and their race.__

b. How many observations are there in the data frame?
```{r}
dim(dat)
```

__There are 6594 observations in the data frame.__

c. Look at the names of the variables in the data frame. Describe what "body_camera", "flee", and "armed" represent, according to the codebook. Again, only write one sentence (max) per variable.
```{r}
names(dat)
```

__The "body_camera" variable indicates whether or not news reports stated that a police officer that was at the scene of the incident was wearing a body camera and may have recorded parts of the incident. The "flee" variable indicates whether or not news reports have started that the victim was moving away from the officer during the incident. The "armed" variable indicates whether or not the victim was armed with any type of weapon that could have been thought to as possible to cause harm by an officer.__

d. What are three weapons that you are surprised to find in the "armed" variable? Make a table of the values in "armed" to see the options.
```{r}
head(table(dat$armed))
```

__Three weapons that I was surprised to find in the dataset were an air conditioner, microphone, and pen.I feel like an air conditioner and microphone are definitely unusual weapons, and the pen does not seem like it would do anything against an officer's gun.__

### Problem 2 (10 points)

a. Describe the age distribution of the sample. Is this what you would expect to see?
```{r}
hist(dat$age, main = "Age Distribution of the Sample", xlab = "Age", xlim = c(0, 100))
```

__The age distribution of the sample is skewed to the right, with a large number of values between ages 20 and 40. This indicates that most people killed by the police in fatal shootings are young, and that older people between ages 60 and 100 are being killed in fatal police shootings at a much lower rate. This is exactly what I expected to see in the age distribution because younger people, especially in the 20 to 40 year old range, are more likely to have encounters with the police, as well as be healthy and be able to try to talk to or provoke the police and try to run away.__

b. To understand the center of the age distribution, would you use a mean or a median, and why? Find the one you picked.
```{r}
median(dat$age, na.rm = TRUE)
```

__To understand the center of the age distribution, I would use the median because the data is skewed to the right and the median is a more robust measure of skewed data as the mean changes with skew or outliers. The median of the age distribution for this dataset is 35 years old, and I removed the missing values since there are only 308 of them out of 6594 observations and they indicate that the person's age was unknown or missing.__

c. Describe the gender distribution of the sample. Do you find this surprising?
```{r}
table(dat$gender)
counts <- table(dat$gender, useNA = "ifany")
barplot(counts, main = "Gender Distrubution", xlab = "Gender", ylab = "Counts", names=c("None", "Females", "Males"))
```

__The gender distribution is definitely skewed, with 6005 more males being fatally shot by police officers than females. This can also be seen in the barplot, with the majority of respondents being males. There were also 3 unknown values in the data, indicating no gender according to the codebook. I do not find this surprising, however, because the majority of police shootings that we hear about in the news are of males, with the rare occurrence of a female. While females are definitely still being killed by the police, it is happening disproportionately to males which is indicated by both the news and the data.__


### Problem 3 (10 points)

a. How many police officers had a body camera, according to news reports? What proportion is this of all the incidents in the data? Are you surprised that it is so high or low?

```{r}
table(dat$body_camera)
910/6594
```

__According to the news reports, 910 police officers had a body camera, which is about 13.80% of all incidents in the data. I am surprised that this is so low because I would think that after the 2014 killing of Michael Brown and subsequent police shootings, police offices and the public would call for an increased use of body cameras.__

b. In  how many of the incidents was the victim fleeing? What proportion is this of the total number of incidents in the data? Is this what you would expect?
```{r}
table(dat$flee)
1058+845+248 #fleeing
2151/6594
```

__There are 2151 incidents of the victim fleeing, which is about 32.62% of all incidents in the data. This was calculated by including the 491 unknown observations in the "not fleeing" group, as what this category of observations represents was not included in the codebook and cannot be assumed to be fleeing, especially since there is already an "other" group. This is pretty much what I would expect because I would think that the majority of people when interacting with the police would not flee, since it is known that the consequences would probably be worse because of it.__

### Problem 4 (10 points) -  Answer only one of these (a or b).

a. Describe the relationship between the variables "body camera" and "flee" using a stacked barplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the options for "flee", each bar contains information about whether the police officer had a body camera (vertically), and the height along the y-axis shows the frequency of that category).*

*Hint 2: Also, if you are unsure about the syntax for barplot, run ?barplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}
library(ggplot2) 
ggplot(dat, aes(fill=body_camera, y=frequency(body_camera), x=flee)) + 
  geom_bar(position="stack", stat="identity") +
  ggtitle("Stacked Barplot of Body Camera Usage and Victim Fleeing") +
  labs(y="Frequency of Body Camera Usage", x = "How the Victim Fled") +
  labs(fill = "Body Camera Usage")
```

__From the plot, it can be seen that everybody that fled from the police did so more often when the police were not wearing a body camera. It looks like not fleeing and the police wearing a body camera are related, as when more police wore a body camera less people fled (since the not fleeing bar had the most police wearing a body camera). You can also see that the most people fled by using a car, which during those indicidents police were less likely to wear a body camera than to actually have it on. While it is unknown what the first bar represents as it is left out of the codebook, it follows the same pattern of the other bars that the police did not wear body cameras more often than they did.__

b. Describe the relationship between age and race by using a boxplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the race categories and the height along the y-axis is age.* 

*Hint 2: Also, if you are unsure about the syntax for boxplot, run ?boxplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}
boxplot <- ggplot(dat, aes(x=race, y=age)) + 
  geom_boxplot() +
  ggtitle("Boxplot of Race and Age")
boxplot
```

__The relationship between race and age looks pretty even, with the median values of age for each race category between 25 and 37. The range for all ages except White is between 12 and 62, while the White range extends from about 5 to 80. There are a significant number of outliers, however, for the Black, Hispanic, and White categories. The NAs for age were removed, and the unknown race category was plotted first on the graph but cannot be interpreted as it is not in the codebook. From this relationship, I can conclude that on average, all races of a similar age between 25 and 37 are killed at similar rates. While there are some outliers to this, especially for the Black and Hispanic categories, all of the medians of age for each race is almost identical.__

### Extra credit (10 points)

a. What does this code tell us? 

```{r, eval=FALSE}
mydates <- as.Date(dat$date)
head(mydates)
(mydates[length(mydates)] - mydates[1])
```

__The as.Date function in R converts the string of date values into actual dates that functions can be applied to. The head function makes sure that this conversion was done correctly by showing us the first 6 values of the data. The "(mydates[length(mydates)] - mydates[1])" function tells us the time difference between the last observation (which is the entire length of mydates) and the first obeservation (which is "mydates[1]), which ends up being 2458 days.__

b. On Friday, a new report was published that was described as follows by The Guardian: "More than half of US police killings are mislabelled or not reported, study finds." Without reading this article now (due to limited time), why do you think police killings might be mislabelled or underreported?

__I think that police killings may be mislabeled or underreported because police offices do not want to admit to have killing anyone or engaging in illegal or overly forceful activities. I think that there may be disparities between the police reporting shooting and witnesses observing shootings, as police offices may be less likely to do so. I also think that police offices may mislabel the shooting as something that was not their fault so that they do not get attacked by the public for the incident.__

c. Regarding missing values in problem 4, do you see any? If so, do you think that's all that's missing from the data?

__Yes, I saw missing values in problem 4 in the age category. The race and fleeing variable also had unknown values, as it was not indicated by the codebook. I do not think that this is all that is missing from the data, as other variables, such as gender, also had missing or unknown values. I think that going along with part b, some of the data reported was mislabeled or incorrect, and this may be represented by the missing or unknown values in the data.__










