---
title: "Here is some of my work in R Studio!"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Portfolio {.tabset}

<h2>Introduction to Data Science</h2>

## Introduction to Data Science

<h2>Statistical Methods for Political Science</h2>

## Statistical Methods for Political Science

<h2>Statistics for the Social Sciences</h2>

## Statistics for the Social Sciences
### Assignment 1
Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.
```{r}
#Install the dataset package  
#install.packages("datasets")

#load in the library 
library(datasets)
```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?
```{r}
#Load in the dataset
USArrests

#rename the USArrests dataset
dat <- USArrests
```
__It is useful to rename the dataset because it is easier to work with. If the data set had a longer or more complicated name, it would be difficult and time consuming to type out the dataset name every time when you want to perform a function on the data. It also lets you keep track of your work if you make different versions of it so that it is not contaminated by changes that were meant to be on one version and not the other. This will let you replicate the work if changes were made to different versions.__











